Modeling tabular icebergs coupled to an ocean model

### 2017-03-22 JAMES_iceberg_only.tex
Update to James

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetime (due to their large size), during which they drift across large distances, altering regional ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models is unable to represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are composed of Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
%We describe the Lagrangian bonded iceberg model and discuss a number of the technical elements involved in its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg drifting away an idealized ice shelf.} 
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf), splitting of large tabular icebergs, and breaking off smaller pieces from the icebergs. Idealized simulations of calving of a tabular iceberg, its drifting and splitting in two smaller tabular icebergs demonstrate capabilities of the developed framework.
\end{abstract}

@@

Large tabular icebergs \-- pieces of floating ice with horizontal dimensions substantially larger than the vertical dimension \-- calve infrequently ($\sim$ every forty-fifty years) from Antarctic or Greenlandic ice shelves \citep{Jacobs1992}. Observational estimates suggest that over the past 30 years approximately half of Antarctic ice-shelf decay is due to iceberg calving, while the other half occurs through ice-shelf melting \citep{Depoorter2013, Rignot2013}.
% While the majority of calved icebergs being small,
The infrequent tabular icebergs (horizontal extent larger than 5 km) account for more than 90\% of the Southern Hemisphere iceberg mass \citep{Tournadre2016}.
%While the majority of calving icebergs are small (horizontal extent smaller than 1 km), the far less numerous large tabular icebergs (horizontal extent larger than 5 km) account for more than 90$\%$ of Southern Hemisphere iceberg mass \citep{Tournadre2016}.

After calving, icebergs slowly drift away from their origins, often becoming stuck in sea ice, or grounding on bathymetric highs along the Antarctic coast \citep{Lichey2001, Dowdeswell2007}. Large tabular icebergs extend deep into the water column, and have the potential to disrupt ocean circulation patterns for months or even years after calving \citep{Robinson2012, Stern2015}.
The freshwater flux from iceberg melt impacts ocean hydrography around the iceberg, influencing sea-ice production and bottom-water formation \citep{Arrigo2002, Robinson2012, Nicholls2009, Fogwill2016}. Because of their large size, the tabular icebergs have long lifetime during which they drift over long distances injecting meltwater along the way and impacting the Southern Ocean state (e.g. hydrography, sea ice conditions, etc.) far away from their calving origins \citep{Stern2016}. Meltwater injection (and the accompanying upwelling) from icebergs can also influence biological productivity by bringing nutrients to the surface ocean or changing sea ice conditions \citep{Arrigo2002, Vernet2012, Biddle2015}. The increased productivity associated with free-floating tabular icebergs has been linked with local increases in ocean carbon uptake, potentially large enough to be a significant fraction of the Southern Ocean carbon sequestration \citep{Smith2007}.

@@

In recent years, there has been an increased interest in iceberg drift and decay. This surge of interest has been driven by (i) the need to understand polar freshwater cycles in order to create realistic climate forecasts and sea level projections \citep{Silva2006, Shepherd2007, Rignot2013}; and (ii) the increased navigation and exploration activities in high-latitudes iceberg-filled waters in the Arctic \citep{Pizzolato2014, Unger2014, Henderson2016}. The increased interest in icebergs has led to the development of numerical models of iceberg drift and decay \citep{Mountain1980, Bigg1997, Gladstone2001, Kubat2005}, some of which have been included in global General Circulation Models \citep{Martin2010, Marsh2015}. These iceberg drift models treat icebergs as Lagrangian point particles, which are advected by the flow, and melt according to certain parameterizations for icebergs melt. Since icebergs are treated as point particles, iceberg drift models are mostly suitable for modeling icebergs smaller than an ocean grid cell. Consequently, these models have mostly been used to represent icebergs smaller than 3.5~km on a global scale \citep{Jongma2009, Martin2010, Marsh2015}.

Point-particle iceberg drift models are less suitable for modeling larger tabular icebergs, where the size and structure of the iceberg may be an important feature in determining their drift and decay \citep{Stern2016}. They also are not suitable for studying the local effects that icebergs have on the surrounding ocean, or the small scale processes that influence iceberg melt and decay \citep{Wagner2014, Stern2015}. For these reasons, tabular icebergs are currently not represented in the iceberg drift models used as components of climate models, despite accounting for the vast majority of the total Southern Hemisphere iceberg mass \citep{Tournadre2016}. Point-particle iceberg models also do not have any representation of iceberg breakup and calving, which is know to be an important iceberg decay mechanism. 
%Similarly, operational iceberg models are often inadequate for projecting iceberg drift at fine scales \citep{Turnbull2015}.

The goal of this study is to develop a new framework to model all kinds of icebergs where tabular icebergs are explicitly resolved. 
Our new representation of icebergs aims to include the following key properties: (i) icebergs should be able to travel large distances within the ocean, (ii)  icebergs should melt and decay as they drift in the ocean, (iii) icebergs should behave as if they has finite extent (in order to model local effects that icebergs have on the surrounding ocean), and (iv) tabular icebergs should be able to break away from ice shelves or break into smaller pieces. Properties (i) and (ii) are common to point-particle icebergs models, while properties (iii) and (iv) are new to the framework developed in this study.

In order to allow icebergs to travel large distance, it is natural to model the icebergs in a Lagrangian framework (as in the point particle iceberg drift models described above). 
In our model we therefore represent icebergs using Lagrangian elements.
However, in the new framework, icebergs are no longer treated as point particles that interact with the ocean at a point location, but rather icebergs are given structure, so that they interact with the ocean across multiple ocean grid cells. To do this, we represent tabular icebergs as a collection of many Lagrangian elements with finite extent which are `bonded' together by numerical bonds. The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit (see schematic in Figure \ref{fig:Hex_schematic}). 
An advantage of this approach is that we can break the numerical bonds to simulate an iceberg detaching from an ice shelf (e.g.: Figure \ref{fig:Spread_area} and \ref{fig:SST_Collapse}) or an iceberg breaking up into smaller pieces.

The manuscript is organized as follows. Section 2 gives a description of many of the key aspects of the model. Since this model is a new approach to modeling icebergs, we present many of the technical aspects of the model. In section 3 and 4, we demonstrate the capabilities of the model by simulating a tabular iceberg detaching from an idealized ice shelf. In a further simulations we sever some numerical bonds within the tabular iceberg to demonstrate iceberg breaking.

@@

\section{Model description}

%In this section we introduce a new framework for modeling tabular icebergs which are explicitly resolved in the ocean. In this framework, icebergs are given a finite extent so that they can be used to model local iceberg-ocean interactions. This framework also allows icebergs to fracture and break into smaller pieces.

The Lagrangian Bonded Iceberg Model (LBIM) is a Lagrangian particle-based model in that the objects of the model are Lagrangian elements. Each element represents a column of ice that is floating in the ocean, and has a position, velocity, mass, and a set of dimensions, which can evolve in time. The motion of each element is determined by a momentum equation which is solved in the (Lagrangian) reference frame of the element. The elements experience oceanic and atmospheric forces, which are either prescribed, or computed by coupling the LBIM to an ocean/atmosphere model. The LBIM elements also interact with one another and can be bonded together to form larger structures.

@@

The ice elements in the LBIM are given finite extent by assigning to each element a finite surface area and shape, and allowing each element to interact with other elements when the boundaries of the elements overlap. The finite extent of the elements is felt by the ocean by distributing the elements' weight and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice elements. 
To simplify the model, we only use ice elements with hexagonal surface area. In order represent icebergs with other shapes, we introduce numerical bonds which allow multiple ice elements to be bonded together to form larger structures. The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit.  The angular momentum of the elements is not modeled explicitly; instead rotational motion of larger structures emerge as a consequence of bond orientation and collective motion. 
%In this latter case, the LBIM can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

@@

\subsection{Interactive Forces} 
%As described above, our model treats icebergs as rigid icebergs with finite extent, which move through the ocean.
The interactive forces in the LBIM model are used to (i) prevent the ice elements from overlapping and (ii) to connect multiple ice elements so that they move together as a rigid body.
Modeling the collisions and movements of rigid objects precisely, requires very small time steps and precise collision detection algorithms, which are very time consuming. Models using these methods are typically only run for a few days or even a few seconds, and are used to study rapid processes like crack formation or ridging \citep{Hopkins2004, Bassis2013, Rabatel2015}). The tabular iceberg framework presented in this study is developed in order to be used in general circulation models used for climate, and therefore must be run for many years. In order to gain the required computational efficiency, we relax the requirement that icebergs must be perfectly rigid and that ice elements can not overlap. Instead, we model the interactive forces between icebergs using damped elastic forces, which can be calculated more rapidly. %A consequence of this is that the tabular icebergs can bend slightly, and ice elements can overlap slightly. However, these effects are sufficiently small that it

@@

To describe the forces between two elements, we begin by introducing some notation.  Let $\vec{x_{i}}$, $\vec{x_{j}}$ be the positions of elements i and j. The distance between elements i and j is
\begin{equation}
\label{distance}
d_{ij}= | \vec{x_{i}} - \vec{x_{j}}|.
\end{equation}
When calculating the interactive forces between elements, the elements are assumed to be circular. We define the interaction diameter of an element by 
\begin{equation}
\label{Diameter}
D_{i}=2 \sqrt{\frac{A_{i}}{\pi}},
\end{equation}
where $A_{i}$ is the planar surface area of element i. Using this, we define the critical interactive length scale, 
\begin{equation}
L_{ij}=\frac{D_{i} + D_{j}}{2},
\label{eq:Interaction_crit}
\end{equation}
which governs interactions between elements $i$ and $j$. 
Repulsive forces are only applied when $d_{i,j} \textless L_{i,j}$, while for $d_{i,j} \textgreater L_{i,j}$ attractive bonded forces are applied when a bond exists between element $i$ and $j$. 
% The bonded forces are designed such that in the absence of other forces, bonded particles will settle in an equilibrium position where elements are separated by the critical interaction length scale $L_{i,j}$.
% (see schematic in Figure \ref{fig:Interactive_forces}).

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements i and j are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F}_{IA})_{ij}$ on an element i by an element j is given by
\begin{equation}
\label{F_interactione}
(\vec{F}_{IA})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements i and j.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{i,j}-L_{i,j} \bigg) T_{i,j} \vec{r}_{ij},
\end{equation}
where $\vec{r}_{ij}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element i and j, $\kappa_{e}$ is the spring constant, and $T_{i,j}$ is the minimum of the thickness of elements i, j. The interactive forces obey Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum thickness, $T_{i,j}$, is preferred to the average thickness, since this means that for two bonded elements a fixed distance apart, the acceleration due to elastic forces is bounded, even when the thickness of one of the elements approaches zero. 
When $d_{ij} = L_{ij}$, the elastic forces is equal to zero, so that bonded particles will settle in an equilibrium position where elements are separated by the critical interaction length scale $L_{i,j}$.

@@

\subsection{Ocean-ice and ice-ocean coupling}
The LBIM is coupled to the ocean model via a two-way synchronous coupling, meaning that ocean-model fields are passed to the LBIM, and the LBIM fields are passed back to the ocean model at every time step. Passing fields between the two models involves interpolating the fields from the ocean model's Eularian grid onto the LBIM's `Lagrangian grid' (i.e.: onto the ice elements, Figure  \ref{fig:Hex_schematic}), and aggregating fields from the Lagrangian elements onto the ocean-model's Eularian grid.

@@

%More care needs to be taken when passing fields back from the LBIM to the ocean model. 
The coupling from the LBIM to ocean model is more complex.
The LBIM influences the ocean by: (i) applying a pressure to the ocean surface, (ii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice, and (iii) imposing heat, salt and mass fluxes on the ocean, associated with ice melting.
Six fields are passed from the LBIM to the ocean model: ice mass, ice area, frictional velocity, heat, salt, and mass fluxes. 
Fields from the LBIM are aggregated from the Lagrangian elements to the Eulerian ocean grid before they are passed to the ocean model.

>>>>

\begin{abstract}
Large tabular icebergs calved from Antarctic ice shelves have long lifetimes (due to their large size), during which they drift across large distances, altering regional ocean circulation, bottom-water formation, sea-ice formation and biological primary productivity in the icebergs' vicinity.
However, despite their importance, the current generation of ocean circulation models is unable to represent large tabular icebergs.
In this study we develop a novel framework to model large tabular icebergs submerged in the ocean. In this framework, tabular icebergs are composed of Lagrangian elements that drift in the ocean, and are held together and interact with each other via bonds.
%We describe the Lagrangian bonded iceberg model and discuss a number of the technical elements involved in its development. 
%We demonstrate the capabilities of the model by modeling a tabular iceberg drifting away an idealized ice shelf.} 
A break of these bonds allows the model to emulate calving events (i.e. detachment of a tabular iceberg from an ice shelf), and tabular icebergs breaking up into smaller pieces. Idealized simulations of a calving tabular iceberg, its drift, and of a tabular iceberg breaking into two pieces, demonstrate capabilities of the developed framework.
\end{abstract}

@@

Large tabular icebergs \-- pieces of floating ice with horizontal dimensions substantially larger than the vertical dimension \-- calve infrequently ($\sim$ every forty-fifty years) from Antarctic or Greenlandic ice shelves \citep{Jacobs1992}. Observational estimates suggest that over the past 30 years approximately half of Antarctic ice-shelf decay is due to iceberg calving, while the other half occurs through ice-shelf melting \citep{Depoorter2013, Rignot2013}.
% While the majority of calved icebergs being small,
The infrequently-calved tabular icebergs (horizontal extent larger than 5 km) account for more than 90\% of the Southern Hemisphere iceberg mass \citep{Tournadre2016}.
%While the majority of calving icebergs are small (horizontal extent smaller than 1 km), the far less numerous large tabular icebergs (horizontal extent larger than 5 km) account for more than 90$\%$ of Southern Hemisphere iceberg mass \citep{Tournadre2016}.

After calving, icebergs slowly drift away from their origins, often becoming stuck in sea ice, or grounding on bathymetric highs along the Antarctic coast \citep{Lichey2001, Dowdeswell2007}. Large tabular icebergs extend deep into the water column, and have the potential to disrupt ocean circulation patterns for months or even years after calving \citep{Robinson2012, Stern2015}.
The freshwater flux from iceberg melt impacts ocean hydrography around the iceberg, influencing sea-ice production and bottom-water formation \citep{Arrigo2002, Robinson2012, Nicholls2009, Fogwill2016}. Because of their large size, the tabular icebergs have long lifetimes during which they drift over long distances injecting meltwater along the way and impacting the Southern Ocean state (e.g. hydrography, sea ice conditions, etc.) far away from their calving origins \citep{Stern2016}. Meltwater injection (and the accompanying upwelling) from icebergs can also influence biological productivity by bringing nutrients to the surface ocean or changing sea ice conditions \citep{Arrigo2002, Vernet2012, Biddle2015}. The increased productivity associated with free-floating tabular icebergs has been linked with local increases in ocean carbon uptake, potentially large enough to be a significant fraction of the Southern Ocean carbon sequestration \citep{Smith2007}.

@@

In recent years, there has been an increased interest in iceberg drift and decay. This surge of interest has been driven by (i) the need to understand polar freshwater cycles in order to create realistic climate forecasts and sea level projections \citep{Silva2006, Shepherd2007, Rignot2013}; and (ii) the increased navigation and exploration activities in high-latitude iceberg-filled waters in the Arctic \citep{Pizzolato2014, Unger2014, Henderson2016}. The increased interest in icebergs has led to the development of numerical models of iceberg drift and decay \citep{Mountain1980, Bigg1997, Gladstone2001, Kubat2005}, some of which have been included in global General Circulation Models \citep{Martin2010, Marsh2015}. These iceberg drift models treat icebergs as Lagrangian point particles, which are advected by the flow, and melt according to certain parameterizations for icebergs melt. Since icebergs are treated as point particles, iceberg drift models are mostly suitable for modeling icebergs smaller than an ocean grid cell. Consequently, these models have mostly been used to represent icebergs smaller than 3.5~km on a global scale \citep{Jongma2009, Martin2010, Marsh2015}.

Point-particle iceberg drift models are less suitable for modeling larger tabular icebergs, where the size and structure of the iceberg may be an important feature in determining their drift and decay \citep{Stern2016}. They also are not suitable for studying the local effects that icebergs have on the surrounding ocean, or the small scale processes that influence iceberg melt and decay \citep{Wagner2014, Stern2015}. For these reasons, tabular icebergs are currently not represented in the iceberg drift models used as components of climate models, despite accounting for the vast majority of the total Southern Hemisphere iceberg mass \citep{Tournadre2016}. Point-particle iceberg models also do not have any representation of iceberg breakup and calving, which is know to be an important iceberg decay mechanism and influences iceberg trajectories.
%Similarly, operational iceberg models are often inadequate for projecting iceberg drift at fine scales \citep{Turnbull2015}.

The goal of this study is to develop a new framework to model all kinds of icebergs, where tabular icebergs are explicitly resolved in the ocean. Our new representation of icebergs aims to include the following key properties: (i) icebergs should be able to travel large distances within the ocean, (ii)  icebergs should melt and decay as they drift in the ocean, (iii) icebergs should behave as if they has finite extent (so that we can use the model to study local effects that icebergs have on the surrounding ocean), and (iv) tabular icebergs should be able to break away from ice shelves or break into smaller pieces. Properties (i) and (ii) are common to point-particle icebergs models, while properties (iii) and (iv) are new to the framework developed in this study. A further requirement of the new framework is that the model should run sufficiently quickly to be used in general circulation models used for climate.

In order to allow icebergs to travel large distances, it is natural to model the icebergs in a Lagrangian framework (as in the point particle iceberg drift models described above). In our model we therefore represent icebergs using Lagrangian elements. However in our model icebergs are no longer treated as point particles that interact with the ocean at a single location, but rather icebergs are given physical structure, so that they interact with the ocean across multiple ocean grid cells, depress the ocean surface over a wide area, and can interact with other icebergs  (see schematic in Figure \ref{fig:Hex_schematic}). This is done by assigning a finite surface area and shape to the Lagrangian elements, which allows the elements to behave as if they has a finite extent. 
%Element interact with the ocean by distributing their weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element.
The finite extent of an element is transmitted by the ocean by distributing the element's weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element. The finite extent of an element is felt by other elements through repulsive forces which are applied when the boundaries of the elements overlap.
%Elements interact with other elements when the boundaries of the elements overlap (see schematic in Figure \ref{fig:Hex_schematic}).

Large tabular icebergs can then be represented by `bonding' together multiple ice elements using numerical bonds (see schematic in Figure \ref{fig:Hex_schematic}).
The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit. 
This allows us to study how tabular icebergs drift in the ocean when forced by ocean currents and wind.
%Figure \ref{fig:Hex_schematic}) shows a schematic of the difference between point-particle icebergs, finite extent icebergs and tabular icebergs constructed out of multiple elements. 
%Figure \ref{fig:Hex_schematic} shows a schematic of point-particle elements, elements with finite extent and bonded elements, and shows schematically how tabular icebergs can be constructed out of multiple elements.
An advantage of representing tabular icebergs using numerical bonds is that by manually breaking the bonds, we can simulate iceberg calving, allowing us to study the ocean response to tabular icebergs detaching from ice shelves (e.g.: Figure \ref{fig:SST_Collapse}), or the response to an iceberg fracturing into multiple smaller pieces.

%The goal of this study is to develop a new framework to model all kinds of icebergs where tabular icebergs are explicitly resolved. Our new representation of icebergs aims to include the following key properties: (i) icebergs should be able to travel large distances within the ocean, (ii)  icebergs should melt and decay as they drift in the ocean, (iii) icebergs should behave as if they has finite extent (so that we can use the model to study local effects that icebergs have on the surrounding ocean), and (iv) tabular icebergs should be able to break away from ice shelves or break into smaller pieces. Properties (i) and (ii) are common to point-particle icebergs models, while properties (iii) and (iv) are new to the framework developed in this study. A further requirement of the new framework is that the model should run sufficiently quickly to be used in general circulation models used for climate.

%In order to allow icebergs to travel large distances, it is natural to model the icebergs in a Lagrangian framework (as in the point particle iceberg drift models described above). In our model we therefore represent icebergs using Lagrangian elements.However, in our framework, icebergs are no longer treated as point particles that interact with the ocean at a point location, but rather each iceberg is given a finite physical structure, which is explicitly resolved in the ocean. This is done by giving each element a finite extent, and using multiple elements to construct larger icebergs. The ice elements are given finite extent by assigning to each element a finite surface area and shape (see schematic in Figure \ref{fig:Hex_schematic}). The finite extent of an element is transmitted by the ocean by distributing the element's weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element. The finite extent of an element is felt by other elements through repulsive forces which are applied when the boundaries of the elements overlap.
%(regular hexagons are used)

%In order represent large tabular icebergs with complex shapes, we introduce numerical bonds that allow multiple ice elements to be bonded together to form larger structures (Figure \ref{fig:Hex_schematic}). The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit (e.g.: Figure \ref{fig:SST_Collapse}). An advantage of this approach is that we can then break the numerical bonds to simulate an iceberg detaching from an ice shelf or an iceberg breaking up into smaller pieces.

%The iceberg model presented in this study is developed in order to be used a component of general circulation models used for climate. This means that the model must (i) be sufficiently computationally efficient to run for multiple years coupled to an ocean model, and (ii) must run on multiple processors. To aid in the computational efficiency, we do not treat the icebergs as fully rigid objects, but instead relax the rigidity requirement and allow elements to overlap and bonds of tabular icebergs to have some flexibility. This approximation allows us to avoid using  complex collision detection algorithms, adaptive time step refinements or even precise calculation of angular momentum budgets, which can be very computationally expensive. Discrete element models using these methods are typically only run for a few days or even a few seconds, and often use elements which are 1 m in length \citep{Hopkins2004, Bassis2013, Rabatel2015}.  In our model, typical size of ice elements is of order 1 km, and typical time steps must be similar to those used for large scale ocean modeling.  In the model description below, we present a number of  techniques used to make the model run faster (time stepping scheme, efficient bond interactions). We also discuss methods which allow the model to be run on multiple processors.

%The iceberg model presented in this study is developed in order to be used a component of general circulation models used for climate. As such, the numerical tools used are limited to those which allow the model to be sufficiently computationally efficient to run for multiple years while coupled to an ocean model.

The manuscript is organized as follows. 
Section 2 gives a description of many of the key aspects of the model developed in this study. Since this model is a new approach to modeling icebergs, we present many of the technical aspects of the model.  In Sections 3 and 4, we demonstrate the capabilities of the model by simulating a tabular iceberg detaching from an idealized ice shelf. In a further simulation we break some numerical bonds within the tabular iceberg to demonstrate an iceberg splitting in two.

@@

\section{Model description}
%In this section we introduce a framework for modeling tabular icebergs which are explicitly resolved in the ocean.
%The goal of this study is to develop a new framework to model all kinds of icebergs where tabular icebergs are explicitly resolved. 
%Our new representation of icebergs aims to include the following key properties: (i) icebergs should be able to travel large distances within the ocean, (ii)  icebergs should melt and decay as they drift in the ocean, (iii) icebergs should behave as if they has finite extent (so that we can use the model to study local effects that icebergs have on the surrounding ocean), and (iv) tabular icebergs should be able to break away from ice shelves or break into smaller pieces. Properties (i) and (ii) are common to point-particle icebergs models, while properties (iii) and (iv) are new to the framework developed in this study. A further requirement of the new framework is that the model should run sufficiently quickly to be used in general circulation models used for climate.

%In order to allow icebergs to travel large distances, it is natural to model the icebergs in a Lagrangian framework (as in the point particle iceberg drift models described above). In our model we therefore represent icebergs using Lagrangian elements. However, in our framework, icebergs are no longer treated as point particles that interact with the ocean at a point location, but rather each iceberg is given a finite physical structure, which is explicitly resolved in the ocean. This is done by giving each element a finite extent, and using multiple elements to construct larger icebergs.

%The ice elements are given finite extent by assigning to each element a finite surface area and shape (see schematic in Figure \ref{fig:Hex_schematic}). The finite extent of an element is transmitted by the ocean by distributing the element's weight, surface area and melt fluxes over multiple ocean grid cells in a way which is consistent with the shape of the ice element. The finite extent of an element is felt by other elements through repulsive forces which are applied when the boundaries of the elements overlap.
%(regular hexagons are used)
%In order represent large tabular icebergs with complex shapes, we introduce numerical bonds that allow multiple ice elements to be bonded together to form larger structures (Figure \ref{fig:Hex_schematic}). The numerical bonds hold the ice elements together and allow a collection of elements to move as a unit (e.g.: Figure \ref{fig:SST_Collapse}). An advantage of this approach is that we can then break the numerical bonds to simulate an iceberg detaching from an ice shelf or an iceberg breaking up into smaller pieces.

The Lagrangian Bonded Iceberg Model (LBIM) is a Lagrangian particle-based model in that the objects of the model are Lagrangian elements. Each element represents a column of ice that is floating in the ocean, and has a position, velocity, mass, and a set of dimensions, which can evolve in time. The motion of each element is determined by a momentum equation which is solved in the (Lagrangian) reference frame of the element. The elements experience oceanic and atmospheric forces, which are either prescribed, or computed by coupling the LBIM to an ocean/atmosphere model. The LBIM elements also interact with one another and can be bonded together to form larger structures. The angular momentum of the elements is not modeled explicitly; instead rotational motion of larger structures emerge as a consequence of bond orientation and collective motion.

@@

%In this latter case, the LBIM can be categorized as a discrete element model (DEM). 
% in the sense that large structures are constructed from smaller independent Lagrangian elements.

@@

\subsection{Interactive Forces} 
%As described above, our model treats icebergs as rigid icebergs with finite extent, which move through the ocean.
The interactive forces in the LBIM model are used to (i) prevent the ice elements from overlapping and (ii) to connect multiple ice elements together so that the collection of elements moves as a rigid body.
Modeling the collisions and movements of rigid objects precisely, requires very small time steps and precise collision detection algorithms, which are very computationally expensive. Models using these methods are typically only run for a few days or even a few seconds, and are used to study rapid processes like crack formation or ridging \citep{Hopkins2004, Bassis2013, Rabatel2015}). The tabular iceberg framework presented in this study is developed in order to be used in general circulation models used for climate, and therefore must be run for many years. In order to gain the required computational efficiency, we relax the requirement that icebergs must be perfectly rigid and that ice elements can not overlap. Instead, we model the interactive forces between ice elements using damped elastic forces, which can be calculated more rapidly. %A consequence of this is that the tabular icebergs can bend slightly, and ice elements can overlap slightly. However, these effects are sufficiently small that it

@@

To describe the forces between two elements, we begin by introducing some notation.  Let $\vec{x_{i}}$, $\vec{x_{j}}$ be the positions of elements $i$ and $j$. The distance between elements $i$ and $j$ is
\begin{equation}
\label{distance}
d_{ij}= | \vec{x_{i}} - \vec{x_{j}}|.
\end{equation}
When calculating the interactive forces between elements, the elements are assumed to be circular. We define the interaction radius of an element by 
\begin{equation}
\label{Diameter}
R_{i}=\sqrt{\frac{A_{i}}{\pi}},
\end{equation}
where $A_{i}$ is the planar surface area of element $i$. Using this, we define the critical-interactive-length scale, 
\begin{equation}
L_{ij}=R_{i} + R_{j},
\label{eq:Interaction_crit}
\end{equation}
which governs interactions between elements $i$ and $j$. 
Repulsive forces are only applied when $d_{i,j} \textless L_{i,j}$, while for $d_{i,j} \textgreater L_{i,j}$ attractive bonded forces are applied when a bond exists between element $i$ and $j$ (see diagram in Figure \ref{fig:Force_diagram}). 
The interactive forces are designed such that (in the absence of other external forces) bonded particles will settle in an equilibrium position where elements are separated by $L_{i,j}$.

% The bonded forces are designed such that in the absence of other forces, bonded particles will settle in an equilibrium position where elements are separated by the critical interaction length scale $L_{i,j}$.
% (see schematic in Figure \ref{fig:Interactive_forces}).

To aid in notation, we define a bond matrix $B_{ij}$ such that $B_{ij}=1$ if elements $i$ and $j$ are bonded together and  $B_{ij}=0$ otherwise. Using this notation, the interactive force $(\vec{F}_{IA})_{ij}$ on an element $i$ by an element $j$ is given by
\begin{equation}
\label{F_interactione}
(\vec{F}_{IA})_{ij}= \left\{%
  \begin{array}{ll}
   (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
    & \mbox{if $d_{ij} \leq L_{ij} $ }\\
 %     & \mbox{if $|x_{i} -x_{j}| \leq \frac{1}{2}(D_{i}+D_{j}) $ }\\
    (\vec{F}_{e})_{ij} + (\vec{F}_{d})_{ij}
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=1 $ }\\
       & \mbox{if $d_{ij} >  L_{ij}  \textrm{ and } B_{ij}=1 $ }\\
      0
 %     & \mbox{if $|x_{i} -x_{j}| > \frac{1}{2}(D_{i}+D_{j}) \textrm{ and } B_{ij}=0 $ }.\\
      & \mbox{if $d_{ij} > L_{ij} \textrm{ and } B_{ij}=0 $ }.\\
  \end{array}
\right.
\end{equation}
$(\vec{F_{e}})_{ij}$ and $(\vec{F_{d}})_{ij}$ are the elastic and frictional damping components of the interactive force between elements $i$ and $j$.
The elastic force $(\vec{F_{e}})_{ij}$ between elements is given by
\begin{equation}
\label{F_elastic}
(\vec{F_{e}})_{ij}= -\kappa_{e}  \bigg( d_{i,j}-L_{i,j} \bigg) T_{i,j} \vec{r}_{ij},
\end{equation}
where $\vec{r}_{ij}= \frac{(\vec{x_{i}} -\vec{x_{j}})}{|\vec{x_{i}} -\vec{x_{j}}|}$ is the directional unit vector between the position of element $i$ and $j$, $\kappa_{e}$ is the spring constant, and $T_{i,j}$ is the minimum of the thickness of elements $i$ and $j$. The interactive forces obey Newton's 3rd Law (i.e.: $(\vec{F}_{IA})_{ij}=-(\vec{F}_{IA})_{ji}$).
%This is in contrast to some other DEM models in the literature (e.g. \citet{Li2014}). 
The minimum thickness, $T_{i,j}$, is preferred to the average thickness, since this means that for two bonded elements a fixed distance apart, the acceleration due to elastic forces is bounded, even when the thickness of one of the elements approaches zero.

@@

\subsection{Ocean-ice and ice-ocean coupling}
The LBIM is coupled to the ocean model via a two-way synchronous coupling, meaning that ocean-model fields are passed to the LBIM, and the LBIM fields are passed back to the ocean model at every time step. Passing fields between the two models involves interpolating the fields from the ocean model's Eularian grid onto the LBIM's `Lagrangian grid' (i.e.: onto the ice elements), and aggregating fields from the Lagrangian elements onto the ocean-model's Eularian grid.

@@

%More care needs to be taken when passing fields back from the LBIM to the ocean model. 
%The coupling from the LBIM to ocean model is more complex.
The LBIM influences the ocean by: (i) applying a pressure to the ocean surface, (ii) affecting the upper ocean by applying a no-slip boundary condition and frictional velocity beneath the ice, and (iii) imposing heat, salt and mass fluxes on the ocean, associated with ice melting.
Six fields are passed from the LBIM to the ocean model: ice mass, ice area, frictional velocity, and heat, salt and mass fluxes. 
Fields from the LBIM are aggregated from the Lagrangian elements to the Eulerian ocean grid before they are passed to the ocean model.

The aggregation of the LBIM fields onto the ocean grid is done in a way that is consistent with the shape of the elements in the LBIM (see Section 2.3). Fields are `spread' to the ocean model grid by exactly calculating what fraction of an element's surface area lies in a particular grid box, and dividing the field in proportion to this fraction.
%While the position of a particular element might be inside one ocean grid cell, the implied shape of the element will often intersect other ocean grid cells. 
As an example, consider a hexagonal element in the LBIM, which is positioned such that it intersects four ocean grid cells (Figure \ref{fig:Hex_intersections}b). In this situation, the element's mass (for example) is divided between these four ocean cells in proportion to the overlap area between the hexagonal element and the grid cell (this fraction is shown by the colors in Figure \ref{fig:Hex_intersections}b). An advantage of this approach is that there are no jumps in pressure as an element moves from one grid cell to another, which could trigger artificial tsunamis within the ocean model, making the ocean model unrealistic. 
%This prevents sudden changes to the aggregated pressure field pressure which can trigger tsunamis within the ocean model, making the ocean model unrealistic.  
%The aggregation scheme is therefore designed such that there are no jumps in pressure as an element moves from one grid cell to the next, and total pressure is conserved as an element moves.
